---
title: "R Review"
output: html_notebook
---

For this review, we'll be working with dispatch data from the Lincoln Police Department. You can find the data online here: [http://opendata.lincoln.ne.gov/datasets/lpd-dispatch-records](http://opendata.lincoln.ne.gov/datasets/lpd-dispatch-records).

Use the Download menu to select a csv download.

Then move the data file you downloaded into your project folder.

### Load packages

Load the tidyverse, janitor and lubridate packages.

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
```




### Load data

Load your data into an object called 'dispatches'.

```{r}
dispatches <- read.csv("LPD_Dispatch_Records.csv")
```

#### Fix column names

Use janitor to make all column names comply with R preferred style - all lowercase and underscores between words.

```{r}
dispatches <- dispatches %>% clean_names()
```



### Analysis questions

#### Datatypes

Look at the documentation for the data. Do all the columns appear to be formatted correctly?

Use lubridate to change the RPT_Date column to ymd format.

```{r}
dates <- dispatches %>% mutate(
    CleanDate = ymd(rpt_date))
```


Find the 13 rows that failed to parse. Why?

How many cases are in the data? How many unique cases?

```{r}
nrow(dispatches)


```
```{r}
UniqueCases <- dispatches %>%
  group_by(cfs_legend) %>%
  summarise(
    total = n()
  )
```
There were 43 different types of cases.
#### Arranging

What are the oldest and most recent cases in the data?

```{r}
  NewCases <- dispatches %>% arrange(desc(rpt_date))
```
The newest case was June 8, 2020

The oldest case was January 1, 2017

#### Filtering

Create a dataframe called 'missing' with just missing persons cases in it.

```{r}
missing <- NewCases %>% filter(rpt_date == "0")
```

Use the str_detect function to find all the cases that mention O Street in the address.


```{r}
OStreet <- dispatches %>% filter(str_detect(block_address, "O ST", negate = FALSE)) 
```



#### Counting

Use the count() function to find the number of dispatches to each neighborhood. 

```{r}
Neighborhood <- dispatches %>%
  group_by(neighbor) %>%
  summarise(
    count = n()
  ) %>% arrange(desc(count))
```

Which neighborhood appears the most in the data?
Downtown

Do you see any limitations of this data when we are counting by neighborhood?

Yes, a lot of the neighborhoods are not specific and cover a wide range of areas

#### Mutating

Create a new column called 'year' that includes just the year from date_fixed.

```{r}
Year <- dates %>%
  mutate(
    year = as.numeric(format(CleanDate,'%Y'))
  )
```



Use the case_when function to create a new categorical variable dividing the rpt_time column into the 24 hours of the day. Be careful to make sure that each time fits into only one value of the new variable.

```{r}
hours <- dates %>%
  mutate(
    hour_of_day = substr(rpt_time, 1, 2)
  )
```


What hour of the day sees the most police action?

```{r}
hours %>%
  group_by(hour_of_day) %>%
  summarise(
    total_per_hour = n()
  ) %>% arrange(desc(total_per_hour))
```

3-5 pm is not a good time to be out and about.
#### Grouping and summarizing 

How many drug-related dispatches occurred in each year of the data?

```{r}
Drugs <- Year %>%
  group_by(year, cfs_legend) %>%
  summarise(
    count = n()
  ) %>% arrange(desc(count))
```
3,300 in 2017. 3,436 in 2018 3,185 in 2019 and 1,252 in 2020

Create a new column called month. Then using group_by and summarize, find the maximum, minimum and average number of dispatches per month.

```{r}
month <- dates %>%
  mutate(
    month = as.numeric(format(CleanDate,'%m'))
  )
```

```{r}
month %>%
  group_by(month) %>%
  summarise(
    total_per_month = n()
  ) %>% arrange(desc(total_per_month))
```
```{r}
monthly <- month %>%
  group_by(month) %>%
  summarise(
    total_per_month = n()
  ) %>% arrange(desc(total_per_month))
```
```{r}
monthlydispatch <- monthly %>% 
  group_by(month) %>% 
  summarise(
    avgdispatch = mean(`total_per_month`), 
    mindispatch = min(`total_per_month`),
    maxdispatch = max(`total_per_month`))
```

 



#### Percent change

What was the percent change in total number of dispatches from 2018 to 2019?


```{r}
YearTotal <-Year %>%
  group_by(year) %>%
  summarise(
    total_per_year = n()
  ) %>% arrange(desc(total_per_year))
```

```{r}
PercentChange <- YearTotal %>% pivot_wider(names_from = year, values_from = total_per_year)
```


```{r}
YearDiff <- PercentChange %>% mutate(
  change = ((`2019` - `2018`)/`2018`*100
) )

```

There was a .8% change from 2018 to 2019
#### Line charts

Using ggplot, create a line chart of the number of cases per month and year. Choose an appropriate color for the line, add a title and labels and a choose a theme.

```{r}
month %>% group_by(CleanDate) %>% summarize(total=n()) %>% ggplot() + geom_line(aes(x=CleanDate, y=total)) +
  labs(x="Date", y="Dispatches", title= "Crimes in Lincoln Since Jan. 2017", subtitle= "The summer looks like a hot spot for crime.", caption="Source: Lincoln Police Department | Graphic: Tim  Hofmann") + 
  theme_minimal() + theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 8), 
    plot.subtitle = element_text(size=10)
  ) +
  ggsave("LPDdispatch.png", width = 8, height = 4)

```


What do you observe about the yearly pattern of police dispatches? Why do you suppose that is?

#### Column charts

Using ggplot, create a column chart that shows the five most common categories in the cfs_legend column. Apply appropriate design decisions to your chart.

```{r}
type <- dispatches %>%
  group_by(cfs_legend) %>%
  summarise(total = n()) %>%
  arrange(desc(total)) %>%
  filter(total >= 22193)
```
```{r}
library(scales)
```


```{r}
ggplot(type, aes(x=reorder(cfs_legend, -`total`), weight=`total`)) + 
  geom_bar() + 
  scale_y_continuous(labels=comma) + 
  labs(x="", y="", title="5 Most Crimes in Lincoln", subtitle= "If your other category is that high, you should probably specify.", caption="Source: Lincoln Police Department | Graphic: Tim Hofmann") + 
  coord_flip() + 
  theme_minimal() + theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 8), 
    plot.subtitle = element_text(size=10)
  ) +
  ggsave("5Dispatches.png", width = 8, height = 4)
```

